Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-06-12 Referee1_Response.tex

\begin{itemize}
\item[16.] \textit{p17, l55: Perhaps add a comment on why DB1 may not experience the asymptotic variance reduction? 
(Is it because it may lack a unique solution?)}
\end{itemize}

\noindent
Done.
% xxx Yes, it is partially because we are not sure if it has a unique optimal solution as DB1 has never been solved before. 
%Even if it has been solved, the variance reduction is shown asymptotically but the computations are only for small sample sizes. 
%We reworded this to reflect these points.
%Thank you for this suggestion. 
\medskip

\begin{itemize}
\item[17.] \textit{p18, figures: The figures are a bit small and hence can be hard to read without enlarging them (but perhaps this will be taken care of at production).}
\end{itemize}

\noindent 
We split these two figures into three and enlarged each graph. 
All the new graphs added to the paper in Section 5.4 (Computational Efficiency) are also larger than the old graphs. 
\medskip

@@

\begin{itemize}
\item[18.] \textit{p18, l57: The authors speak of coverage probabilities for DB1, but how can this be done without knowing an optimal solution to DB1?}
\end{itemize}

\noindent 
We used a Latin hypercube sample of size 50,000 estimate $\mu_{\hat{x}}$ for DB1, which was then used to estimate the coverage probability of the OMRP CIs for this test problem.
We explain this at the end of Section 5.1 (Test Problems). 
We also changed the wording in Section 5.3 (Results of Experiments) to make it clear that the coverage probability is estimated for DB1. 
\medskip

@@

\begin{itemize}
\item[19.] \textit{p19: It may be possible to expand on the last paragraph of Section 5, in giving guidance on selecting the degree of overlap, depending on the authors' response to comment M1 above.}
\end{itemize}

\noindent 
Thank you for this suggestion. 
xxx We expanded this paragraph based on our findings. 
xxx Write more---hopefully, we will find good stuff!!!
\medskip

### 2014-06-12 omrp_lb_rev.tex
Enlarged graphs, changed comp results and relevant referee responses

\begin{proof} When the sampling problems used in the calculation of $\gbb$ all have optimal solutions that solve (SP), i.e., when $\xs_{j} = \xs$, for all $j=1,2,\ldots, \nbl$, then $\gbb_l = \dbb_l$. 
Therefore, complement of this event provides an upper bound on the probability that $\dbb_l - \gbb_l \neq 0$. 
Let $E^C$ denote the complement of event $E$. 
Using this upper bound and continuing on, we obtain
	\begin{align}
		\p{\dbb_l - \gbb_l \neq 0} & \leq \p{\left( \bigcap_{j=1}^{\nbl} \{\xs_{j} = \xs\} \right)^C} \notag \\
		& = \p{ \bigcup_{j=1}^{\nbl} \{\xs_{j} \neq \xs\}} \notag \\
% 		& \leq \sum_{j=1}^{\nbl} \p{ \xs_j \neq \xs } \notag \\
		& \leq \left( \nbl\right) \p{ \xs_{m_l} \neq \xs }. \notag
	\end{align}
	With a batching structure in which $m_l = \lfloor n_l^{r} \rfloor$ for some $0<r<1$,  $\gamma_l \geq 1$, and by A3, the desired result follows.
\end{proof}

@@

A summary of results from these experiments is depicted in Figures~\ref{fig:nv} and \ref{fig:varvar}.
Figures~\ref{fig:nv}(a) and \ref{fig:varvar} show the reduction of variance, with each term being normalized with respect to $\var{nVG}$ of the nonoverlapping case (MRP). 
The solid lines in these figures show the theoretical variance reduction from \citep{Welch1987} given by (\ref{eq:var_reduct_formula}). 
Empirical results agree with the theoretical values well, with more variability observed in CEP1, 10D, and DB1. 
These experiments suggest that similar variance reduction can be achieved in stochastic programming by overlapping the batches, even with small sample sizes.
We note that the variance reduction in DB1 can be less than the asymptotic amount. %for the small sample sizes used in these experiments. 
Due its size, DB1 has not yet been solved to optimality, therefore, it may lack a unique optimal solution. 
Even with the relatively small sample sizes compared to the size of this problem, however, overlapping appears to achieve some variance reduction.

\begin{figure}[htb!]
	\centering
	\subfloat[][Variance Reduction]{
	   \label{fig:nv_varvar}
	   \includegraphics[width=.32\textwidth]{images/nv_varvar_discrete}}
	\subfloat[][Expected Variance]{
	   \label{fig:nv_avgvar}
	   \includegraphics[width=.32\textwidth]{images/nv_avgvar_discrete}}
	\subfloat[][Coverage Probability]{
	   \label{fig:nv_cover}
	   \includegraphics[width=.32\textwidth]{images/nv_cover_discrete}}
	\caption{
		Summary of results for NVD. 
		(a) Reduction in variance of $nVG$,
		(b) estimates of $\e{nVG}$, and 
		(c) coverage probability of the CIs for various values of $\gamma/m$ ($\gamma/m=1$ denotes the nonoverlapping batches).
	}
\label{fig:nv}
\end{figure}

\begin{figure}[htb!]
	\centering
	\subfloat[][CEP1]{
	   \label{fig:cep1_varvar}
	   \includegraphics[width=.32\textwidth]{images/cep1_varvar}}
	\subfloat[][10D]{
	   \label{fig:10d_varvar}
	   \includegraphics[width=.32\textwidth]{images/10D_varvar}}
	\subfloat[][DB1]{
	   \label{fig:db1_varvar}
	   \includegraphics[width=.32\textwidth]{images/db1_varvar}}
		\caption{
		Variance Reduction for
		(a) CEP1,
		(b) 10D, and
		(c) DB1
		for various values of $\gamma / m$.
		}
\label{fig:varvar}
\end{figure}

Figure~\ref{fig:nv}(b) shows estimates of $\e{nVG}$ in the newsvendor problem for changing batch size and amount of overlap. 
We can see that the expectation of the variance estimator is not changed with increasing overlap. 
Although the estimate of $\e{nVG}$ looks large for $m=100$, the error from $\sigma^2_\xh$ is never more than 1\%.  
Estimates for $\e{nVG}$ showed a similar pattern for CEP1, 10D, and DB1.  
Those graphs are omitted for brevity.

Finally, Figure~\ref{fig:nv}(c) shows the coverage probability of CIs generated by OMRP for several values of $m$ across varying values of $\gamma/m$ for the newsvendor problem.  
The results of the classical overlapping batches estimators show that coverage probability does not change with the amount of overlap, which we can  empirically see in this figure for OMRP for NVD.
We observed the same for CEP1, 10D, and DB1.  
The coverage probabilities from applying the (nonoverlapping) MRP algorithm to these problems presented in \citep{Bayraksan2006} agree with our results: for the newsvendor problem, coverage probability drops as $m$ increases and the coverage probability of CEP1 (not shown) remains fairly constant around the desired value of $90\%$. 
Coverage probability for 10D (not shown) and {\it estimated} coverage probability for DB1 (also not shown) were very high, more than $0.99$.  
The bias from solving (\ref{eq:sto_prog_m}) for these problems seems to be much more significant than the variance.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}
The authors thank Andrzej Ruszczy{\'{n}}ski and Artur {\'{S}}wietanowski for
access to their regularized decomposition code, which was used to solve CEP1 and DB1. 
%This research is supported in part by the National Science Foundation under Grants DMS-0602173 and EFRI-0835930.  
An earlier version of this paper appeared in \citep{love2011overlapping}.

>>>>

\begin{proof} When the sampling problems used in the calculation of $\gbb$ all have optimal solutions that solve (SP), i.e., when $\xs_{j} = \xs$, for all $j=1,2,\ldots, \nbl$, then $\gbb_l = \dbb_l$. 
Therefore, probability of the complement of this event provides an upper bound on the probability that $\dbb_l - \gbb_l \neq 0$. 
Let $E^C$ denote the complement of event $E$. 
Using this upper bound and continuing on, we obtain
	\begin{align}
		\p{\dbb_l - \gbb_l \neq 0} & \leq \p{\left( \bigcap_{j=1}^{\nbl} \{\xs_{j} = \xs\} \right)^C} \notag \\
		& = \p{ \bigcup_{j=1}^{\nbl} \{\xs_{j} \neq \xs\}} \notag \\
% 		& \leq \sum_{j=1}^{\nbl} \p{ \xs_j \neq \xs } \notag \\
		& \leq \left( \nbl\right) \p{ \xs_{m_l} \neq \xs }. \notag
	\end{align}
	With a batching structure in which $m_l = \lfloor n_l^{r} \rfloor$ for some $0<r<1$,  $\gamma_l \geq 1$, and by A3, the desired result follows.
\end{proof}

@@

A summary of results from these experiments is depicted in Figures~\ref{fig:varvar1}--\ref{fig:nv}.
Figures~\ref{fig:varvar1} and \ref{fig:varvar2} show the reduction of variance, with each term being normalized with respect to $\var{nVG}$ of the nonoverlapping case (MRP). 
The solid lines in these figures show the theoretical variance reduction from \citep{Welch1987} given by (\ref{eq:var_reduct_formula}). 
Empirical results agree with the theoretical values well, with more variability observed in CEP1, 10D, and DB1. 
These experiments suggest that similar variance reduction can be achieved in stochastic programming by overlapping the batches, even with small sample sizes.
We note that the variance reduction in DB1 can be less than the asymptotic amount. %for the small sample sizes used in these experiments. 
Due its size, DB1 has not yet been solved to optimality, therefore, it may lack a unique optimal solution. 
Even with the relatively small sample sizes compared to the size of this problem, however, overlapping appears to achieve some variance reduction.

\begin{figure}[htb!]
	\centering
	\subfloat[][NVD]{
	   \label{fig:nv_varvar}
	   \includegraphics[width=.5\textwidth]{images/nv_varvar_discrete}}
	\subfloat[][CEP1]{
	   \label{fig:cep1_varvar}
	   \includegraphics[width=.5\textwidth]{images/cep1_varvar}}
	\caption{
		$\var{n VG}$ relative to the nonoverlapping counterpart across various values of $\gamma/m$ ($\gamma/m=1$ denotes the nonoverlapping batches) for test problem
		(a) NVD, and
		(b) CEP1
	}
\label{fig:varvar1}
\end{figure}

\begin{figure}[htb!]
	\centering
	\subfloat[][10D]{
	   \label{fig:10d_varvar}
	   \includegraphics[width=.5\textwidth]{images/10D_varvar}}
	\subfloat[][DB1]{
	   \label{fig:db1_varvar}
	   \includegraphics[width=.5\textwidth]{images/db1_varvar}}
		\caption{
		$\var{n VG}$ relative to the nonoverlapping counterpart across various values of $\gamma/m$ ($\gamma/m=1$ denotes the nonoverlapping batches) for test problem
		(a) 10D, and
		(b) DB1
		}
\label{fig:varvar2}
\end{figure}

\begin{figure}[htb!]
	\centering
	\subfloat[][Expected Variance]{
	   \label{fig:nv_avgvar}
	   \includegraphics[width=.5\textwidth]{images/nv_avgvar_discrete}}
	\subfloat[][Coverage Probability]{
	   \label{fig:nv_cover}
	   \includegraphics[width=.5\textwidth]{images/nv_cover_discrete}}
	\caption{ 
		Estimates of
		(a) $\e{nVG}$, and 
		(c) coverage probability of the CIs for various values of $\gamma/m$ ($\gamma/m=1$ denotes the nonoverlapping batches)
		 for NVD
	}
\label{fig:nv}
\end{figure}

Figure~\ref{fig:nv}(a) shows estimates of $\e{nVG}$ in the newsvendor problem for changing batch size and amount of overlap. 
We can see that the expectation of the variance estimator is not changed with increasing overlap. 
Although the estimate of $\e{nVG}$ looks large for $m=100$, the error from $\sigma^2_\xh$ is never more than 1\%.  
Estimates for $\e{nVG}$ showed a similar pattern for CEP1, 10D, and DB1.  
Those graphs are omitted for brevity.

Finally, Figure~\ref{fig:nv}(b) shows the coverage probability of CIs generated by OMRP for several values of $m$ across varying values of $\gamma/m$ for the newsvendor problem.  
The results of the classical overlapping batches estimators show that coverage probability does not change with the amount of overlap, which we can  empirically see in this figure for OMRP for NVD.
We observed the same for CEP1, 10D, and DB1.  
The coverage probabilities from applying the (nonoverlapping) MRP algorithm to these problems presented in \citep{Bayraksan2006} agree with our results: for the newsvendor problem, coverage probability drops as $m$ increases and the coverage probability of CEP1 (not shown) remains fairly constant around the desired value of $90\%$. 
Coverage probability for 10D (not shown) and {\it estimated} coverage probability for DB1 (also not shown) were very high, more than $0.99$.  
The bias from solving (\ref{eq:sto_prog_m}) for these problems seems to be much more significant than the variance.

\subsection{Computational Efficiency}
\label{ssec:compeff}

xxx Add new graphs here \bigskip

xxx Write new text, etc. \bigskip

xxx Revise the recommendations on gammabar \bigskip

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}
The authors thank Andrzej Ruszczy{\'{n}}ski and Artur {\'{S}}wietanowski for
access to their regularized decomposition code, which was used to solve CEP1 and DB1. 
The authors are also grateful to the associate editor and two anonymous referees for valuable suggestions that significantly improved both the content and the presentation of the paper. 
%This research has been supported in part by the National Science Foundation under Grant(s)  EFRI-0835930 and CMMI-1345626 ??? 
An earlier version of this paper appeared in \citep{love2011overlapping}.
